---
title: "A1. Math Review"  
author: "Merdyanto"  
# date: "`r Sys.Date()`"  
output:   
  pdf_document:  
    latex_engine: xelatex
    highlight: haddock  
    keep_tex: true  
    includes:
      in_header: latex-topmatter_pdf.tex
    toc: True   
    toc_depth: 2
    # number_sections: true  
monofont: Consolas
smaller: yes
classoption: a4paper
header-includes: 
  - \usepackage{tikz}
  - \usepackage{pgfplots}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage 

# I. Differentiation and integration

## Exercise 1

Suppose $f(x)=xe^x$, find $f'(x)$. 

*\underline{Solution:}* 

\vspace{-20pt}

\begin{eqnarray}
&& f(x)=xe^x \nonumber\\
&& \Longrightarrow f'(x)=(x)'e^x+x(e^x)' \nonumber\\
&& \Longrightarrow f'(x)=e^x+xe^x \nonumber
\end{eqnarray}

\hr

## Exercise 2

Suppose $f(x)=e^{2x}$, find $f'(x)$.

*\underline{Solution:}* To use the theorem of composite function, the function $f()$ can be seen as $f(x)=h(g(x))$, where $h(x)=e^x$ and $g(x)=2x$. It follows $f'(x)=e^{2x} \times 2 = 2e^{2x}$.

\hr

## Exercise 3

Derive $\int{f'(x)g(x) dx} = f(x)g(x) - \int{f(x)g'(x) dx}.$ *(Hint: Use Theorem 2 above.)*

*\underline{Solution:}* From the theorem, we have $(f(x)·g(x))'=f'(x)g(x)+f(x)g'(x)$. Applying antiderivative to both hand sides gives

\begin{equation}
f(x)\cdot g(x) = \int f'(x)g(x)dx + \int f(x)g'(x)dx
\end{equation}

It follows $\int f'(x)g(x)dx=f(x)g(x)-\int f(x)g'(x)dx$.

\newpage

## Exercise 4 

Find $\int{xe^xdx}$,and evaluate $\int_{0}^{1}{xe^x}dx$. *(Hint: Use Exercise 3 above.)*

*\underline{Solution:}* \
*\underline{Basic Formula:}* \
apply integration by parts: u = x, v' = e^x.\
$e^xx-\int e^xdx$.\
if it follow $\int e^xx - e^x$\
$e^xx - e^x$\
finally add constant to the equation.\
$e^xx - e^x + c$\

*\underline{Using above theorem to solve exercise 4}* \
Integral $\int_{0}^{1} {xe^x}dx$\
Following above equation basic formula\
$((e^1)1 - e^1)-((e^0)0 - e^0)$\
substitute\
$((e - e)-((1)0 - 1)$\
$((0)-(0 - 1)$\
1\

\newpage

# II. Numerical methods for finding a root

## Exercise 5

\lc

The original R code:

\vspace{10pt}

```{r}
f <- function(x) {
  return(1+1/x)
}
tol <- 10^(-5)
x_now <- 0.1
repeat{
  x_next <- f(x_now)
  if (abs(x_next-x_now) < tol) {
    break
  }
  x_now <- x_next
  print(x_next)
}
```

\rc

The comparable python code:

\vspace{10pt}

```{phyton}
def f(x):
    return 1+(1/x)
tol = 10^(-5)
x_now = 0.1

while x_next := f(x_now):
    if abs(x_next - x_now)<tol:
        break
    x_now = x_next
    print(x_next)
```
\ec

\newpage

# III. Matrix algebra

## Exercise 6

*\underline{Solution:}*

$$
\begin{pmatrix} 
.6 & .4 \\
\end{pmatrix}
\begin{pmatrix}
.7 & .3\\ 
.5 & .5
\end{pmatrix} =
(0.6 \cdot 0.7+0.4 \cdot 0.5 \quad 0.6 \cdot 0.4+0.4 \cdot 0.5) = (0.62 \quad 0.38)
$$

\hr

## Exercise 7

What is $P^2$?

*\underline{Solution:}*

$$
P^2=
\begin{pmatrix} 
0.7 & 0.3 \\ 
0.5 & 0.5
\end{pmatrix}
\begin{pmatrix}
0.7 & 0.3\\ 
0.5 & 0.5
\end{pmatrix} = 
\begin{pmatrix}
0.7 \cdot 0.7 + 0.3 \cdot 0.5 & 0.7 \cdot 0.3 +0.3 \cdot 0.5\\ 
0.5 \cdot 0.7 + 0.5 \cdot 0.5 & 0.5 \cdot 0.3 +0.5 \cdot 0.5
\end{pmatrix} \\
$$
$$
= \begin{pmatrix}
0.49+0.15  & 0.21+0.15\\ 
0.35+0.25 & 0.15+0.25
\end{pmatrix} \\
$$

$$
= \begin{pmatrix}
0.64 & 0.36\\ 
0.6 & 0.4
\end{pmatrix}
$$ 

\hr

\newpage

## Exercise 8

\begin{eqnarray}
\begin{pmatrix} 
\mbf{v}_{1}&\mbf{v}_{2}
\end{pmatrix}
\begin{pmatrix} 
.7&.3 \\
.5&.5
\end{pmatrix}&=&
\begin{pmatrix}
\mbf{v}_{1} & \mbf{v}_{2}
\end{pmatrix} \nonumber\\
\mbf{v}_{1}+\mbf{v}_{2} &=& 1 \nonumber
\end{eqnarray}

*\underline{Solution:}*
*\underline{Equation 1}* \

\begin{eqnarray}
\begin{pmatrix} 
\mbf{v}_{1} \cdot 0.7 + \mbf{v}_{2} \cdot 0.5 & \mbf{v}_{1} \cdot 0.3 + \mbf{v}_{2} \cdot 0.5  
\end{pmatrix}
\begin{pmatrix} 
.7&.3 \\
.5&.5
\end{pmatrix}&=&
\begin{pmatrix}
\mbf{v}_{1} & \mbf{v}_{2}
\end{pmatrix} 
\end{eqnarray}\
*\underline{Equation 2}* \

\begin{eqnarray}
\mbf{v}{1}+\mbf{v}{2} &=& 1 \nonumber
\end{eqnarray}

Calculate equation 1 became two part\
1.${0.7\mbf{v}_{1} + 0.5\mbf{v}_{2}}= \mbf{v}_{1}$\
2. combine with equation 2\
3.$\mbf{v}_{2} = 0.375$\

from above equation return to equation 2\
\begin{eqnarray}
\mbf{v}{1}+\mbf{v}{2} &=& 1 \nonumber
\end{eqnarray}\
\begin{eqnarray}
\mbf{v}{1} &=& 0.625 \nonumber
\end{eqnarray}

\newpage

## Exercise 9

Solve the following system of equations.

\begin{eqnarray}
x &=& y \\
y &=& 0.5z \\
z &=& 0.6-0.4x \\
x+y+z &=& 1
\end{eqnarray}

*\underline{Solution:}*
*\underline{Equation 1}*
\begin{eqnarray}
x &=& y \\
\end{eqnarray}
*\underline{Equation 2}*
\begin{eqnarray}
y &=& 0.5z \\
\end{eqnarray}
*\underline{Equation 3}*
\begin{eqnarray}
z &=& 0.6-0.4x \\
\end{eqnarray}
*\underline{Equation 4}*
\begin{eqnarray}
x+y+z &=& 1 \\
\end{eqnarray}
*\underline{Combining equation 1, equation 2, and equation 4}*\
\begin{eqnarray}
y+y+z &=& 1 \\
2y + z &=& 1 \\
2(0.5)+z &=& 1 \\
z+z &=& 1 \\
z &=& 0.5 \\
\end{eqnarray}
*\underline{Above equation became equation 5}*\

enter equation 5 to equation equation 2\
\begin{eqnarray}
y &=& 0.5z \\
y &=& 0.25 \\
\end{eqnarray}\
above equation became equation 6\
enter equation 6 to equation 1\
\begin{eqnarray}
x &=& y \\
x &=& 0.25 \\
\end{eqnarray}\

\hr
\newpage

## Exercise 10

$$
\begin{pmatrix} \mbf{v}_0 & \mbf{v}_1 & \mbf{v}_2 \end{pmatrix}
\begin{pmatrix} -2 & 2 \\ 3 & -5 & 2 \\ & 3 &-3 \end{pmatrix} =
\begin{pmatrix} 0 & 0 & 0 \end{pmatrix}
$$
$$ \mbf{v}_0 + \mbf{v}_1 + \mbf{v}_2 = 1 $$
*\underline{Solution:}* 

$$
\begin{pmatrix} \mbf{v}_0 & \mbf{v}_1 & \mbf{v}_2 \end{pmatrix}
\begin{pmatrix} -2 & 2 & 0 \\ 3 & -5 & 2 \\0 & 3 &-3 \end{pmatrix} =
\begin{pmatrix} 0 & 0 & 0 \end{pmatrix}
$$
assume
if we break down after we solve the equation of the matrices we get equation 1,2, and 3\
and we assume $\mbf{v}_0 = x$, $\mbf{v}_1 = y$, $\mbf{v}_2 = z$ to simply the math\
\begin{eqnarray}
-2x + 3y &=& 0 \\
2x+ 5y + 3z &=& 0 \\
2y+3z &=& 0 \\
\end{eqnarray}

we use substitution-elimination of equation (28) to the addition matrices\
the result became
\begin{eqnarray}
y &=& 0.26 \\
\end{eqnarray}

then we substitute y to the equation 27 we got 
\begin{eqnarray}
x &=& 0.39 \\
\end{eqnarray}

finally we can get z using the addition matrix equation\
\begin{eqnarray}
z &=& 0.35 \\
\end{eqnarray}

in conclusion to revert back\
$\mbf{v}_0 = 0.39$, $\mbf{v}_1 = 0.26$, $\mbf{v}_2 = 0.35$

\newpage

## Exercise 11

**Solve the following system of equations.**

$$
\begin{pmatrix} \mbf{v}_1 & \mbf{v}_2 & \mbf{v}_3 & \mbf{v}_4 \end{pmatrix}
\begin{pmatrix} .7 & .3 &0&0 \\ .5 & .5&0&0 \\ 0&0 & .6 & .4 \\ 0&0& .3 & .7 \end{pmatrix} =
\begin{pmatrix} \mbf{v}_1 & \mbf{v}_2 & \mbf{v}_3 & \mbf{v}_4 \end{pmatrix}
$$

$$\mbf{v}_1 + \mbf{v}_2 + \mbf{v}_3 + \mbf{v}_4 = 1$$
$$\mbf{v}_1 + \mbf{v}_2 = a$$

*\underline{Solution:}* 

$$
\begin{pmatrix}
0.7\cdot \mbf{v}_1 + 0.5\cdot \mbf{v}_2 & 0.3\cdot \mbf{v}_1+0.5\cdot \mbf{v}_2 & 0.6\cdot \mbf{v}_3 + 0.3\cdot \mbf{v}_4& 0.4\cdot \mbf{v}_3+0.7\cdot \mbf{v}_4 
\end{pmatrix}=
\begin{pmatrix} \mbf{v}_1 & \mbf{v}_2 & \mbf{v}_3 & \mbf{v}_4 \end{pmatrix}
$$
and we can assume $\mbf{v}_1 = m$, $\mbf{v}_2 = n$, $\mbf{v}_3 = o$ , $\mbf{v}_4 = p$ to simply the math\
\begin{eqnarray}
0.7m + 0.5n &=& m \\
0.3m+ 0.5n &=& n \\
0.6o+0.3p &=& o \\
0.4o+0.7p &=& p
\end{eqnarray}
equation 37 and equation 38 will produce\
\begin{eqnarray}
0.7m + 0.5n &=& m \\
n &=& \frac{3}{5}n \\
\end{eqnarray}
equation 39 and equation 40 will produce\
\begin{eqnarray}
0.6o + 0.3p &=& o \\
p &=& \frac{4}{3}o \\
\end{eqnarray}
therefore the addition matrices became\
\begin{eqnarray}
m+n+o+p &=& 1 \\
\frac{8}{5}m + \frac{7}{3}o &=& 1 \\
o &=& \frac{3}{7}(1-\frac{8}{5}m) \\
o &=& \frac{3}{7} -\frac{24}{35}m \\
\end{eqnarray}

substitute equation 50 to the equation 45\
\begin{eqnarray}
p &=& \frac{4}{7} - \frac{32}{35}m \\
\end{eqnarray}

using the addition matrices and substite o and p we can get m\
\begin{eqnarray}
m &=& \frac{5}{8} \\
\end{eqnarray}

if we can get m then we can get n\
\begin{eqnarray}
n &=& \frac{3}{8} \\
\end{eqnarray}

therefore 
\begin{eqnarray}
a &=& \frac{5}{8}+\frac{3}{8} \\
a &=& 1
\end{eqnarray}



\newpage

## Exercise 12

Solve following and express $\mbf{v}_i$ for $i=0,1,2,...$

\begin{eqnarray}
\mbf{v}_0 + \mbf{v}_1 + \mbf{v}_2 +... &=& 1 \nonumber\\
0.02 \mbf{v}_0 + 0.02 \mbf{v}_1 + 0.02 \mbf{v}_2 + ... &=& \mbf{v}_0 \nonumber\\
0.98 \mbf{v}_0 &=& \mbf{v}_1 \nonumber\\
0.98 \mbf{v}_1 &=& \mbf{v}_2 \nonumber\\
0.98 \mbf{v}_2 &=& \mbf{v}_3 \nonumber\\
...&=&... \nonumber
\end{eqnarray}

*\underline{Solution:}* 
using summation notation behaviour we can assume\
\begin{eqnarray}
\sum_{i=0}^{v}i=1
\end{eqnarray}

also the notation behaviour we can get\
\begin{eqnarray}
0.02 \times \sum_{i=0}^{v}i = {v}_0
\end{eqnarray}
\begin{eqnarray}
\sum_{i=0}^{v}i = 50\times{v}_0
\end{eqnarray}
from equation 60 and 62 we can get\
\begin{eqnarray}
{v}_0 = 0.02
\end{eqnarray}

substitute equation 63 and we can get the question answer\
\begin{eqnarray}
0.98 \times 0.02 = {v}_1
{v}_1 = 0.0196
\end{eqnarray}

and the rest question\
\begin{eqnarray}
{v}_2 = \frac{49}{125000}
\end{eqnarray}
\begin{eqnarray}
{v}_3 = 7.89 \times 10^-6
\end{eqnarray}
\hr