---
title: "A2. Prob Review"
author: "Merdyanto"
output:
  pdf_document:
    latex_engine: xelatex
    highlight: haddock
    keep_tex: yes
    includes:
      in_header: latex-topmatter_pdf.tex
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
monofont: Consolas
smaller: yes
classoption: a4paper
header-includes:
- \usepackage{tikz}
- \usepackage{pgfplots}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage 

# I. Probability

## Exercise 1

*Show that* $\mathbb{P}(A \mid B \cap C)\mathbb{P}(B \mid C)=\mathbb{P}(A \cap B \mid C)$

*\underline{Solution:}* 

given\
$\mathbb{P}(A \mid B)= \frac{\mathbb{P}(A \cap B)}{\mathbb{P}(B)}$


\begin{eqnarray}
\mathbb{P}(A \mid B \cap C)\mathbb{P}(B \mid C) 
  &=& \frac{\mathbb{P}(A \cap (B \cap C)}{\mathbb{P}(B \cap C)} \times \frac{\mathbb{P}(A \cap B)}{\mathbb{P}(C)}
\end{eqnarray}

\begin{eqnarray}
 &=& \frac{\mathbb{P}(A \cap B \cap C)}{\mathbb{P}(C)}\
\end{eqnarray}

\begin{eqnarray}
&=& \mathbb{P}(A \cap B \mid C)
\end{eqnarray}


# II. Random variables

\newpage

# III. Uniform distribution

## Exercise 2

$X \sim U(10,20)$, then what is $F(10)$? and $F(15)$?

*\underline{Solution:}*

\[f(x)=
\begin{Bmatrix}
\frac{1}{10} \,\,(10\leq x\leq 20) \\ \\
0 \,\,\,\,\,\,\,\,\,\,\,\,\, otherwise
\end{Bmatrix}
\]
\vspace{10pt}
\[F(x)=
\begin{Bmatrix}
0 \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,(10\leq x) \\\\ \frac{x-10}{10} \,\,\,\,\,\,\,(10\leq x \leq 20) \\\\1 \,\,\,\,\,\,\,\,\,\,\,\,\,(x>20)
\end{Bmatrix}
\]

$F(10) = \frac{x-10}{10}$\
$F(10) = \frac{10-10}{10}$\
$F(10) = 0$\

$F(15) = \frac{x-10}{10}$\
$F(15) = \frac{15-10}{10}$\
$F(15) = \frac{1}{2}$\

\newpage

# IV. Exponential distribution

## Exercise 3

Prove that pdf $\rightarrow$ cdf*

*\underline{Solution:}*

$F(x)=\int_{-\infty}^{x}f(x)\,\,dx$

i) if $x<0$, then $f(x)=0$, so $F(x)=0$
ii) if $x \ge 0$, then 

\begin{eqnarray}
F(x) &=&  P(X <=x)
\end{eqnarray}

\begin{eqnarray}
&=& \int_{-\infty}^{\infty}f(x)\,\,dx
\end{eqnarray}

\begin{eqnarray}
&=& \int_{-\infty}^{0}f(x)\,\,dx + \int_{0}^{x}f(x)\,\,dx
\end{eqnarray}

\begin{eqnarray}
&=& 0 + \int_{0}^{x}f(x)\,\,dx
\end{eqnarray}

\begin{eqnarray}
&=& f(x) - f(0)\\
&=& 0\\
\end{eqnarray}

\newpage
\hr

## Exercise 4

Show that $\mathbb{E}X=1/\lambda$.

*\underline{Solution:}*
for $\mathbb{E}X=1/\lambda$

to solve above equation, use integration partial where 
\begin{eqnarray}
\int u  du &=& uv - \int v dv \nonumber\\
u &=&  \lambda x \nonumber\\
dv &=&  e^{-\lambda x} \nonumber
\end{eqnarray}


\begin{eqnarray}
\mathbb{E}X 
  &=& \int_{-\infty}^{\infty} x f(x) \,\, dx \\
  &=& \int_{-\infty}^{0} x \lambda e^{-\lambda x}  \,\, dx + \int_{0}^{\infty} x \lambda e^{-\lambda x}  \,\, dx \\
  &=& 0 +  \int_{0}^{\infty} x \lambda e^{-\lambda x}  \,\, dx \\
\end{eqnarray}


\begin{eqnarray}
&=& \left.- \lambda x \cdot \frac{1}{\lambda} e^{-\lambda x}\right|_0^\infty - \frac{1}{\lambda} e^{-\lambda x} dx
\end{eqnarray}

know that
\begin{eqnarray}
\lim_{x\to \infty} \frac{1}{\lambda e^{\lambda x}} \nonumber
&=& 0 \nonumber
\end{eqnarray}

\begin{eqnarray}
&=& - \left.x e^{-\lambda x} \right|_0^\infty + \int_{0}^{\infty} x \lambda e^{-\lambda x}  \,\, dx \\
&=& - (0 - 0) + \frac{\lambda}{\lambda} \int_{0}^{\infty} x \lambda e^{-\lambda x}  \,\, dx \\
&=& \frac {1}{\lambda}
\end{eqnarray}

\newpage

## Exercise 5 

For $X \sim exp(\lambda)$, show that $Var(X) = 1/\lambda^2$ (Hint: You need to do $EX^2$ first)

*\underline{Solution:}*

Recall that $Var(X)=\E X^2-(\E X)^2$. We already have that $\E X = 1/\lambda$, so it is only left to find an expression for $\E X^2$ in terms of $\lambda$. This can be done in the following way:

\begin{eqnarray}
M_x(t) &=& \int_{0}^{\infty} e^{tx} \lambda e^{-\lambda x}  \,\, dx \\
&=& \lambda \int_{0}^{\infty} e^{tx - \lambda x}  \,\, dx \\
&=& \lambda \int_{0}^{\infty} e^{tx - \lambda x}  \,\, dx \cdot \frac{\lambda - t}{\lambda - t}\\
&=& \frac{\lambda}{\lambda - t} \int_{0}^{\infty} {\lambda - t} e^{tx - \lambda x}  \,\, dx\\
&=& \frac{\lambda}{\lambda - t}
\end{eqnarray}

using double differential equation\

\begin{eqnarray}
\E X^2 &=& \int_{0}{\infty} x^2 \lambda e^{-\lambda x} \,\, dx\\
&=& \left.\frac{d^2}{dt^2} M_x(t)\right|_{t=0} \\ 
&=& \frac{d}{dt} \left.\frac{\lambda}{{\lambda - t}^2} \right|_{t=0} \\
&=&  \frac{-2\lambda}{{\lambda - t}^3} \cdot -1 \\
&=& \frac{-2}{\lambda ^2}
\end{eqnarray}



\newpage

## Exercise 6

Prove the previous theorem

*\underline{Solution:}*

Claim, $\mathbb{P}(X>s+t \mid X>t)=\mathbb{P}(X > s)$.

\begin{eqnarray}
\mathbb{P}(X>s+t \mid X>t) 
&=& \frac{\mathbb{P}(X>s+t \mid X>t)}{\mathbb{P}(X > s)} 
\end{eqnarray}

we know that

\begin{eqnarray}
\mathbb{P}(X>x) &=& 1 - \mathbb{P}(X = x) \nonumber\\
&=& 1 - (1 - e^{-ex}) \nonumber\\
&=& e^{-ex} \nonumber
\end{eqnarray}

\begin{eqnarray}
&=& \frac{\mathbb{P}(X>s+t)}{\mathbb{P}(X > s)} \\
&=& \frac{e^{-\lambda {s+t}}}{e^{-xs}} \\
&=& e^{-\lambda s - \lambda t + \lambda s} \\
&=& e^{-\lambda t} \\
&=& {P}(X>t) 
\end{eqnarray}

\newpage

# V. Poisson distribution

## Exercise 7

For $X \sim Poi(\lambda)$, prove that $\mathbb{E}X=\lambda$.

*\underline{Solution:}*

\begin{eqnarray}
M_x(t) &=& \E {e^{\lambda x}} \\ 
&=& \sum_{k=0}^{\infty} e ^{tk} \cdot \frac{e^{-\lambda} \lambda^k}{k!} \\ 
&=& e^-\lambda \cdot \sum_{k=0}^{\infty} \frac{\lambda ^k e^{tk}}{k!} \\
&=& e^{-\lambda} \cdot e^{\lambda e^t} \\
&=& e^{x \lambda ^ t - \lambda}
\end{eqnarray}

Using equation (39) and differential equation \
\begin{eqnarray}
\E x &=& \frac{d}{dt} \cdot \left. M_x(t) \right|_{t=0} \\
&=& \left.e^{xe^t-\lambda} \lambda e^t \right|_{t=0} \\
&=& e^{\lambda - \lambda} \lambda \\
&=& \lambda
\end{eqnarray}

\newpage

## Exercise 8

Prove the above theorem.

*\underline{Solution:}*

Let us consider the cdf of $min(X_1,X_2)$.

\begin{eqnarray}
\mathbb{P}\left(min(X_1,X_2) \le x\right)
&=& 1-\mathbb{P}\left(min(X_1,X_2) > x\right) \\
&=& 1 - \mathbb{P} (X_1 > x, X_2 > x) \\
&=& 1 - \mathbb{P} (X_1 > x) \mathbb{P} (X_2 > x) \\ 
&=& 1 - {1 - \mathbb{P} (X_1 \leq x)} {1 - \mathbb{P} (X_2 \leq x)}  \\
&=& 1 - (({1- F_{x1}}) \cdot ({1- F_{x2}}))
\end{eqnarray}

\newpage

# VI. Some exercises

## Exercise 9

For $X \sim U(20,40)$, evaluate $\mathbb{E}[X \wedge 25]$ and $\mathbb{E}[(25-X)^+]$

*\underline{Solution:}*

We have

\[f(x)=
\begin{Bmatrix}
\frac{1}{20} \,\,(20\leq x\leq 40) \\ \\0 \,\,\,\,\,\,\,\,\,\,\,\,\, otherwise
\end{Bmatrix}
\]

\begin{eqnarray}
\mathbb{E}[X \wedge 25]
&=& \mathbb{E}[min(X,25)] 
\end{eqnarray}

\[\mathbb{E}[min(X,25)]=
\begin{Bmatrix}
x \,\,if \ (x\leq 25) \\ \\25 \,\,\, otherwise
\end{Bmatrix}
\]

\begin{eqnarray}
\mathbb{E}[(25-X)^+]
&=& \int_{-\infty}^{\infty} \mathbb{E}[(25-X)^+] f(x) dx \\
&=& \int_{-\infty}^{20} \mathbb{E}[(25-X)^+] f(x) dx + \int_{20}^{40} \mathbb{E}[(25-X)^+] f(x) dx + \int_{40}^{\infty} \mathbb{E}[(25-X)^+] f(x) dx \nonumber\\
&=& 0 + \int_{20}^{40} ((25-X),0) \frac{1}{20} dx + 0 \\
&=& \int_{20}^{25} (25-x,0) \frac{1}{20} dx + \int_{25}^{40} 0 \cdot \frac{1}{20} dx \\
&=& \int_{20}^{25} (25-x) \frac{1}{20} dx \\
&=& \left.(25-x) - (\frac{1}{20} \cdot \frac{1}{2}x^2)\right|_{20}^{25} \\
&=& 1.25 - 5.625 \\
&=& -4.375
\end{eqnarray}

\newpage

## Exercise 10

For $X \sim Poi(8)$, 

1. $\mathbb{P}(X=0)$
2. $\mathbb{P}(2\leq X\leq 4)$
3. $\mathbb{P}(X>2)$

*\underline{Solution:}*

We have $\mathbb{P}(X)=\frac{8^{x}e^{-8}}{x!}$

Number 1.
\begin{eqnarray}
\mathbb{P}(X=0) &=& \frac{\lambda^{x}e^{-\lambda}}{x!} \\
&=& \frac{8^0}{0!}e^{-8} \\
&=& e^{-8}
\end{eqnarray}

Number 2.
\begin{eqnarray}
\mathbb{P}(2\leq X\leq 4) &=& \sum_{i=2}^{4} \mathbb{P}(x=i) \\ 
&=& \sum_{i=2}^{4} \frac{8^{i}e^{-8}}{i!}\\
&=& 0.01 + 0.29 + 0.06 \\
&=& 0.36
\end{eqnarray}

Number 3.
\begin{eqnarray}
\mathbb{P}(X > 2) &=& 1 - \mathbb{P}(X \leq 2) \\
&=& 1 - \sum_{i=0}^{2} \frac{8^{i}e^{-8}}{i!}\\
&=& 1 - 41 e^{8}
\end{eqnarray}


\newpage
\hr

## Exercise 11

*For X~exp(7), evaluate*$\mathbb{E}[max(X,7)].$

*\underline{Solution:}*

Remind that $f(x)=7e^{-7x}$.

\begin{eqnarray}
\mathbb{E}[max(X,7)] 
&=& \int_{-\infty}^{\infty} max(x,7) f(x) dx\\
&=& \int_{-\infty}^{0} max(x,7) f(x) dx + \int_{0}^{\infty} max(x,7) f(x) dx \\
&=& 0 + \int_{0}^{\infty} max(x,7) f(x) dx \\
&=& \int_{0}^{7} max(x,7) 7e^{-7x} dx + \int_{7}^{\infty} max(x,7) 7e^{-7x} dx \\ 
&=& 49 e^{-49} - 49 + \int_{7}^{\infty} x 7e^{-7x} dx \\
&=& 49 (e^{-49}-1) + (7 \cdot \frac{\lambda}{\lambda} \int_{7}^{\infty} x 7e^{-7x} dx) \\
&=& -49 + (\frac{7}{\lambda}  \int_{7}^{\infty} x \lambda 7e^{-7x} dx) \\
&=& -49 + frac{7}{\lambda ^2} \\
&=& 7(\lambda^{-2} - 7)
\end{eqnarray}

\newpage

## Exercise 12

For $X \sim exp(8)$, find $x^*$ such that $F(x^*)=0.6$.

*\underline{Solution:}*

while CDF \
\[f_x(x)=
\begin{Bmatrix}
1 -(\lambda e^{-\lambda x}) \,\,if \ (x\ \geqslant 0) \\ \\0 \,\,\, (x\ < 0) 
\end{Bmatrix}
\]

following CDF function \

\begin{eqnarray}
F(x^*) &=& 1 - \lambda e^{-\lambda x^*} \\
e^{-\lambda x^*} &=& 0.4 \\
-\lambda x^* &=& \log{0.4} \\ 
x^* &=& -\frac{1}{8}\log{0.4} \\
x^* &=& 0.049
\end{eqnarray}
\newpage
\hr

## Exercise 13

For $X \sim U(10,20)$, find $x^*$ such that $F(x^*)=0.7$

*\underline{Solution:}*

Remind 

\[F(x)=
\begin{Bmatrix}
0 \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,(10\leq x) \\ \\ \frac{x-10}{10} \,\,\,\,\,\,\,(10\leq x \leq 20) \\
\\1 \,\,\,\,\,\,\,\,\,\,\,\,\,(x>20)
\end{Bmatrix}
\]

For $F(x^*)$ to be 0.7, $x^*$ should be between 10 and 20. Otherwise, it would be only 0 or 1. It follows $x^*=17$. \
CDF \
\[F(x)=
\begin{Bmatrix}
0 \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,(x \leq 10) \\\\ \frac{x-10}{10} \,\,\,\,\,\,\,(10 < x \leq 20) \\\\1 \,\,\,\,\,\,\,\,\,\,\,\,\,(x>20)
\end{Bmatrix}
\]

so \

\begin{eqnarray}
F(x^*) &=& \frac {x^*-10}{10}
\end{eqnarray}

\begin{eqnarray}
\frac {x^*-10}{10} = 0.7 \\ 
x^*-10 = 7 \\
x^* = 17
\end{eqnarray}


```{r, results='hide'}
"A2_Work"
```
