---
title: "A3 and A4"  
author: "Merdyanto"  
# date: "`r Sys.Date()`"  
output:   
  pdf_document:  
    latex_engine: xelatex
    highlight: haddock  
    keep_tex: true  
    includes:
      in_header: latex-topmatter_pdf.tex
    toc: True   
    toc_depth: 2
    # number_sections: true  
monofont: Consolas
smaller: yes
classoption: a4paper
header-includes: 
  - \usepackage{tikz}
  - \usepackage{pgfplots}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage 

# I. Probability

## Exercise 1.1

is $\bar{X}$ an unbiased estimator for $\mu$, why or why not?\

*\underline{Answer:}* 

Based on per understanding and looking theorem 1\

\begin{eqnarray}
\bar{X} &=& \frac{1}{N-1} \sum_{n=1}^{N} X_i
\end{eqnarray}

\begin{eqnarray}
\mathbb{E}X
&=&\frac{1}{N-1} \sum_{n=1}^{N} X_i \\
&=& \frac{1}{N-1} \cdot N \mu \\
&=& \frac{N \mu}{N-1}
\end{eqnarray}

based on above equation, then it is  *\underline{biased estimator}*  because $\mathbb{E}X \neq \mu$

## Exercise 1.2

is $\bar{X}$ an consistent estimator for $\mu$, why or why not?\

*\underline{Answer:}* \

if the number $N \rightarrow \infty$ so we can safely assume that
\begin{eqnarray}
\frac{N \mu}{N-1} \approx \frac{N \mu}{N}
\end{eqnarray}

then it is *\underline{consistent estimator}* .

## Exercise 1.3

is $\bar{X}$ a maximum likelihood estimator for $\mu$, why or why not?\
based on this\

*\emph{In statistics, maximum likelihood estimation (MLE) is a method of estimating the parameters of a probability distribution by maximizing a likelihood function, so that under the assumed statistical model the observed data is most probable. The point in the parameter space that maximizes the likelihood function is called the maximum likelihood estimate.}*\

therefore it is *\underline{maximum likelihood estimator}* because we are trying to find the optimal value for the mean or starndard deviation for distribution given more than 30 samples measurements.

## Exercise 2

You randomly sample 1,600 Korean male and measured their hand length. The sample  mean is 20cm and the sample standard deviation is 2cm. What is the 95% confidence  interval for Korean male’s hand length? \

*\underline{Answer:}* \

known that $n = 1600$, $\mu = 20$, and $s = 2$ then given 95% confidence interval and since n is bigger than 120  we can use $t_{\alpha} = 1.960$ \

\begin{eqnarray}
\bar{x} \pm t_{\alpha} \frac{s}{\sqrt{n}} &=& 20 \pm 1.96 \frac{20}{\sqrt{2}} \\
&=& 20 \pm 27.718 \\
&=& (-7.718 , 47.718)
\end{eqnarray}

Therefore the confidence interval is $(-7.718 , 47.718)$ .

\newpage

# II. Simulation 1

## Exercise 1

Write the two programs in the previous page with python and compare the computation time. \

*\underline{Answer:}* \

\lc

The vectorized phyton code:

\vspace{10pt}

```{python}
import math
import random
import time 
beg_time = time.time()
random.seed(1234)
n = 0
MC_N= 1000000
for i in range(MC_N):
    x = random.random() 
    y = random.random()
    t = math.sqrt(x*2 + y*2)
    if t <= 1:
        n +=1
pi_hat = 4 * n / MC_N
end_time = time.time()
duration = end_time - beg_time
print('Time gap is:', duration ,'s')
```

\rc

The comparable python code:

\vspace{10pt}

```{python}
import math
import random
import time 

beg_time = time.time()
random.seed(1234)
def pi_hat(MC_N):
    count = 0
    for _ in range(MC_N):
        x = random.uniform(0,1)
        y = random.uniform(0,1)
        t = math.sqrt(x**2 + y**2)
        if t <= 1:
            count += 1
    return 4 * count/MC_N
pi_hat(MC_N = 1000000)
end_time = time.time()
duration = end_time - beg_time
print('Time gap is:', duration ,'s')
```

\ec


\newpage


## Exercise 2

Do the above experiment with MC_N increased by the factor of ten, and present the confidence interval.\

*\underline{Answer:}* \

The original R code:

\vspace{5pt}

```{r}
pi_simulator3 <- function(MC_N) {
  x <- runif(MC_N)*2-1
  y <- runif(MC_N)*2-1
  t <-sqrt(x^2+y^2)
  pi_hat <- 4*sum(t<=1)/MC_N
  return(pi_hat)
}
n <- 100
MC_N <- 1000
set.seed(1234)
samples <- rep(0, n)
for(i in 1:n) {
  samples[i] <- pi_simulator3(MC_N)
}
X_bar <- mean(samples)
s <- sqrt(sum((X_bar-samples)^2)/(n-1))
t <- qt(p=0.975, df = n-1)
lb <- X_bar - t*s/sqrt(n) 
ub  <- X_bar + t*s/sqrt(n)
X_bar
s
t
lb
ub
```

therefore the confidence interval is (3.126709, 3.147291)

\newpage

## Exercise 3

Do the Exercise above with n increased by the factor of ten, and present the confidence interval \

*\underline{Answer:}* \

\vspace{5pt}

```{r}
pi_simulator3 <- function(MC_N) {
  x <- runif(MC_N)*2-1
  y <- runif(MC_N)*2-1
  t <-sqrt(x^2+y^2)
  pi_hat <- 4*sum(t<=1)/MC_N
  return(pi_hat)
}
n <- 100
MC_N <- 1000
set.seed(1234)
samples <- rep(0, n)
for(i in 1:n) {
  samples[i] <- pi_simulator3(MC_N)
}
X_bar <- mean(samples)
s <- sqrt(sum((X_bar-samples)^2)/(n-1))
t <- qt(p=0.975, df = n-1)
lb <- X_bar - t*s/sqrt(n) 
ub  <- X_bar + t*s/sqrt(n)
X_bar
s
t
lb
ub
```

therefore the confidence interval is (3.126709, 3.147291)

\newpage

## Exercise 4


Present a previous page’s table by writing a neat python code block\

*\underline{Answer:}* \

\vspace{5pt}

```{python}
from prettytable import PrettyTable

x = PrettyTable()
x.field_names = ["MC_N", "n", "Length of CI"]
x.add_row([1000,100, 0.020])
x.add_row([10000,100, 0.0066])
x.add_row([10000,1000, 0.00205])
print(x)
```


```{r, results='hide'}
"A3A4_Work"
```
