---
title: "A5 and A6"
author: "Merdyanto"
output:
  pdf_document:
    latex_engine: xelatex
    highlight: haddock
    keep_tex: yes
    includes:
      in_header: latex-topmatter_pdf.tex
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
monofont: Consolas
smaller: yes
classoption: a4paper
header-includes:
- \usepackage{tikz}
- \usepackage{pgfplots}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage 

# I. Inverse Transverse Method

## Exercise 1

Using ruinf() function in R, complete the following code block that generates 1,000 random numbers that follow function exp(3)\

*\underline{Answer:}* \

```{r}
set.seed(1234)
N <- 1000
u <- runif(N)
x <- -log(1-u)/3
head(x)
```

\newpage

## Exercise 2.1 Uniform random numbers

```{python}
import matplotlib.pyplot as plt
import random
random.seed(1234)

x = [random.uniform(-1, 1) for _ in range(1000)]
plt.show(plt.hist(x))
```

\newpage

## Exercise 2.2 Normal random numbers


```{python}
import matplotlib.pyplot as plt
import numpy as np
import random

random.seed(1234)
mu = 3
sigma = 2
n = 1000
x = np.random.normal(mu, sigma, n)
plt.show(plt.hist(x))
```


\newpage

## Exercise 2.3 Exponential random numbers


```{python}
import matplotlib.pyplot as plt
import numpy as np
import random

random.seed(1234)
beta = 1/5
lamda = 1 / beta
n = 10000

x = np.random.exponential(lamda, n)
plt.show(plt.hist(x, bins = 20))
```

\newpage
## Exercise 2.4 Poisson random numbers


```{python}
import matplotlib.pyplot as plt
import numpy as np
import random
random.seed(1234)
beta1 = 1/5
lamda1 = 1 / beta
n1 = 10000
x1 = np.random.poisson(lamda, n)
plt.show(plt.hist(x1, bins = 20))
```

\newpage 


# II. Running Estimate Approach

## Exercise 1

Write a python code that produces results in page 10 and 11 using Eq (3). Use the variable names of new_est, old_est, alpha, MC_tgt, MC_err to demonstrate your understanding. \
*\underline{Answer:}* \
\newpage
\lc
The first phyton code:
\vspace{5pt}
```{python}
import random
import matplotlib.pyplot as plt
import math
random.seed(1234) 
old_est = 0
n = 1
MC_N= 1000000
results = []
for i in range(MC_N):
    x = random.uniform(0,1)
    y = random.uniform(0,1)
    t = math.sqrt(x**2 + y**2)
    if t <= 1:
        t = t +1
    mc_target = 4 * t
    alpha = 1/n
    beta = 1- alpha
    old_est = beta * old_est
    mc_err = mc_target - old_est
    new_est = old_est + alpha * mc_err
    results.append(new_est)
    if n > MC_N:
        break
    n = n+1
    old_est = new_est
plt.plot(results[0:20])
plt.ylabel(' pi approximation')
plt.xlabel('iteration #')
plt.show(plt.xticks(range(0,21)))
```
\rc
The comparable python code:
\vspace{5pt}
```{python}
import random
import matplotlib.pyplot as plt
import math
random.seed(1234) 
old_est = 0
n = 1
MC_N= 1000000
results = []
for i in range(MC_N):
    x = random.uniform(0,1)
    y = random.uniform(0,1)
    t = math.sqrt(x**2 + y**2)
    if t <= 1:
        t = t +1
    mc_target = 4 * t
    alpha = 1/n
    beta = 1- alpha
    old_est = beta * old_est
    mc_err = mc_target - old_est
    new_est = old_est + alpha * mc_err
    results.append(new_est)
    if n > MC_N:
        break
    n = n+1
    old_est = new_est
plt.plot(results[1000:1020])
plt.ylabel(' pi approximation')
plt.xlabel('iteration #')
plt.show(plt.xticks(range(0,21)))
```

\ec
\newpage

```{r, results='hide'}
"A5A6_Work"
```
