---
title: "A7 & A8"
author: "Merdyanto"
output:
  pdf_document:
    latex_engine: xelatex
    highlight: haddock
    keep_tex: yes
    includes:
      in_header: latex-topmatter_pdf.tex
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
monofont: Consolas
smaller: yes
classoption: a4paper
header-includes:
- \usepackage{tikz}
- \usepackage{pgfplots}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

\newpage 

# I. Problem and Solution

## Exercise 1

Assume that D follows the following discrete distribution\

*\underline{Answer:}* \

\begin{center}
\begin{tabular}{ c c c c c }
 d & 20 & 25 & 30 & 35\\ 
 P[D = d] & 0.1 & 0.2 & 0.4 & 0.3\\  
 30 \land \: d & 20 & 25 & 30 & 30\\
 $\mathbb(30-X)^+$ & 10 &5 &0&0 \\
 24 \land \: d & 20 & 24 & 24 & 24\\
 $\mathbb (24-X)^+$ & 4 &0 &0& 0 
\end{tabular}
\end{center}

and with above value, we can use equation 1 to get each of the expectation\

\begin{eqnarray}
\mathbb{E}[x] = \sum P[D = d]. d
\end{eqnarray}

\begin{eqnarray}
\mathbb{E}[30 \wedge D] &=& \mathbb{E}[min(30,D)] \nonumber \\
&=& 29.5
\end{eqnarray}

\begin{eqnarray}
\mathbb{E}[(30-D)^+] &=& \mathbb{E}[max(30-D,0)] \nonumber \\
&=& 1
\end{eqnarray}

\begin{eqnarray}
\mathbb{E}[24 \wedge D] &=& \mathbb{E}[min(24,D)] \nonumber \\
&=& 23.6
\end{eqnarray}

\begin{eqnarray}
\mathbb{E}[(24-D)^+] &=& \mathbb{E}[max(24-D,0)] \nonumber \\
&=& 0.4
\end{eqnarray}

\newpage

# II. Exercises
## Exercise 2

Find your brother's optimal stock level by the above Theorem 2. Then, find his expected profit using the Remark 1.\

*\underline{Answer:}* \

\begin{eqnarray}
F(y) \geq \frac{Cu}{Co + Cu} &=& \frac{1}{0.5 + 1} \nonumber \\
&=& \frac{2}{3}
\end{eqnarray}

Finding the optimal stock level using theorem 2\
The minimum $x^*$ where\

\begin{eqnarray}
\frac{x^* - a}{b-a} \geq \frac{2}{3} \nonumber\\
\frac{x^* - 11}{15-11} \geq \frac{2}{3}\nonumber\\
x^* \geq \frac{41}{3}
\end{eqnarray}

therefore optimal stock level is $\frac{41}{3}$. \ 
then based on the Remark 1 we can get expected value \

\begin{eqnarray}
\mathbb{E}[profit] &=& \mathbb{E}[sales rev] + \mathbb{E}[salvage rev] - \mathbb{E}[material cost]
\end{eqnarray}

using equal probability 20% each of the demand from 11 to 15.\
\begin{eqnarray}
\mathbb{E}[profit] &=& \frac{29}{3} \times 0.2 +  \frac{67}{5} \times 0.2 + \frac{38}{3} \times 0.2 + \frac{41}{3} \times 0.2 + \frac{41}{3} \times 0.2 \nonumber\\
&=& 12.6
\end{eqnarray}

therefore expected profit is $12.6.

\newpage

## Exercise 3

Find optimal stock level and expected profit.\

*\underline{Answer:}* \

Known The customer demands follow U(20,40)\
$p = 2$\
$c = 1$\
$s = 0.5$\
$Cu = p -c = 2 -1 = 1$\
$Co = c-s = 1-0.5 = 0.5$\

\begin{eqnarray}
F(y) &=& \frac{1}{1+0.5} \nonumber\\
&=& \frac{2}{3}
\end{eqnarray}

finding the optimal stock then\
\begin{eqnarray}
\frac{x^* - a}{b-a} \geq \frac{2}{3} \nonumber\\
\frac{x^* - 20}{40-20} \geq \frac{2}{3}\nonumber\\
x^* \geq \frac{100}{3}
\end{eqnarray}

therefore the optimal stock will be level is $\frac{100}{3}$ and expected profit using tabular method is $26.5.
\newpage

## Exercise 4
Express the following quantity using sale as X and demand as D.\

*\underline{Answer:}* \
Known that:\
$p = 18$\
$c = 3$\
$s = 1$\
$Cu = p -c = 18 -3 = 15$\
$Co = c-s = 3-1 = 2$\

\begin{eqnarray}
\mathbb{E}[economic cost] &=& \mathbb{E}[understock cost] + \mathbb{E}[overstock cost] \\
&=& Cu\mathbb{E}[(D-X)^+] + Co\mathbb{E}[(X-D)^+] \nonumber\\
&=& 15\mathbb{E}[(D-X)^+] + 2\mathbb{E}[(X-D)^+]
\end{eqnarray}

In newsvendor problem, maximizing the expected profit is equivalent to minimizing the expected  economic cost. Therefore\
\begin{eqnarray}
\mathbb{E}[Profit] &=& argmin_x \: \mathbb{E}[economic cost]\\
&=& argmin_x \: 15\mathbb{E}[(D-X)^+] + 2\mathbb{E}[(X-D)^+]
\end{eqnarray}

\newpage

## Exercise 5

Prove theorem 1.\

\begin{eqnarray}
\mathbb{E}[Profit] &=& argmin_x \: \mathbb{E}[economic cost] \nonumber\\
&=& argmin_x \: Cu\mathbb{E}[(D-X)^+] + Co\mathbb{E}[(X-D)^+]
\end{eqnarray}

*\underline{Answer:}* \

Derivative of this expression with respect to X.\

\begin{eqnarray}
\frac{d}{dX} \mathbb{E}[economic cost] &=& 
\mathbb{E}[understock cost] + \mathbb{E}[overstock cost] \nonumber\\
&=& p(-Xf(X)) - p + pX F'(X) + pF(X) + c \nonumber\\
&=& pF(X) + c - p
\end{eqnarray}

will result negative of the derivative arrived at above, and this is a minimization instead of a maximization formulation, so the critical point will be the same. Therefore the minimum point will be to maximize the expected profit.
 
\newpage

## Exercise 6

Show that $(D \wedge X) + (Y-D)^+ = Y$ .\

*\underline{Answer:}* \

\begin{eqnarray}
min(D,Y) + max(Y-D,0) &=& Y \nonumber
\end{eqnarray}

\lc
\vspace{10pt}
\[min(D,Y)=
\begin{Bmatrix}
D \,\,if \,(D\leq Y) \\ \\
0 \,\,\,\,\,\,\, otherwise
\end{Bmatrix}
\]
\rc
\vspace{10pt}
\[max(Y-D,0)=
\begin{Bmatrix}
Y-D \,\,if \,(Y-D \ge 0) \\ \\
0 \,\,\,\,\,\,\, otherwise
\end{Bmatrix}
\]
\ec

let D equal to Demand, Y is equal to stock and t is time to meet and therefore we can give profit is P and c is cost\

also known that 
\begin{eqnarray}
min(D,Y) &=& Y - (Y-D)^+
\end{eqnarray}

therefore:\

\begin{eqnarray}
min(D,Y) + max(Y-D,0) &=& Y \nonumber\\
Y - (Y-D)^+ + max(Y-D,0) &=& Y  \nonumber\\
Y - max(Y-D,0) + max(Y-D,0) &=& Y \nonumber\\
Y &=& Y \nonumber
\end{eqnarray}

\newpage

## Exercise 7

Let D be a continous random variable and uniformly distributed between 5 and 10.\

$\mathbb{E}[max(D,8)]$

*\underline{Answer:}* \

\[max(D,8)=
\begin{Bmatrix}
D \,\,if \,(D \ge 8) \\ \\
8 \,\,\,\,\,\,\, otherwise
\end{Bmatrix}
\]

\begin{eqnarray}
\mathbb{E}[max(D,8)] &=& \int_{-\infty}^{\infty} max(D,8) f(D) dD \nonumber\\
&=& \int_{-\infty}^{5} max(D,8) f(D) dD + \int_{5}^{10} max(D,8) f(D) dD + \int_{5}^{\infty} max(D,8) f(D) dD \nonumber\\
&=& 0 + \int_{5}^{10} max(D,8) f(D) dD + 0 \nonumber\\
&=& \int_{5}^{8} 8 \frac{1}{5} dD + \int_{8}^{10} (D) \frac{1}{5} dD \nonumber\\
&=& 12 \nonumber
\end{eqnarray}


$\mathbb{E}[(D,8)^-]$

*\underline{Answer:}* \

$(D,8)^-$ is negative part of (D-8)

\[min(D-8,0)=
\begin{Bmatrix}
D-8 \,\,if \,(D \leq 8) \\ \\
0 \,\,\,\,\,\,\, otherwise
\end{Bmatrix}
\]

\begin{eqnarray}
\mathbb{E}[min(D-8,0)] &=& \int_{-\infty}^{\infty} min(D-8,0) f(D) dD \nonumber\\
&=& \int_{-\infty}^{5} min(D-8,0) f(D) dD + \int_{5}^{10} min(D-8,0) f(D) dD + \int_{5}^{\infty} min(D-8,0) f(D) dD \nonumber\\
&=& 0 + \int_{5}^{10} min(D-8,0) f(D) dD + 0 \nonumber\\
&=& \int_{5}^{8} (D-8) \cdot \frac{1}{5} dD + \int_{8}^{10} 0 \cdot \frac{1}{5} dD \nonumber\\
&=& (19.5 - 4.8) + 0 \nonumber \\
&=& 14.7 \nonumber
\end{eqnarray}


\newpage

# III. Advanced Modelling 1
## Exercise 1

Now assume that the demand for bananas is exponentially distributed with mean 1000.

Identify Co, Cu
*\underline{Answer:}* \

Co  = cost - salvage = 30 - (-5) = 35\
Cu  = profit - cost = 60 - 30 = 30\

How many pound of bananas should David order each week?\

*\underline{Answer:}* \

```{r}
Co <- 35
Cu <- 30
criticalvalue <- Cu/(Cu+Co)
xstar <- -1000 * log(1-criticalvalue)
xstar
```


If David order 800 pounds this week, how much profit he should expect this week?\

*\underline{Answer:}* \

```{r}

X <- 800
D <- dexp(X,rate = 10^-3) 
sales_rev <- 60 * pmax(D,X)
salvage_rev <- -5 * pmin(X-D,0)
material_cost <- -30*X
exp_profit <- mean(sales_rev + salvage_rev + material_cost) 
exp_profit
```



\newpage

# IV. Advanced Modelling 2
## Exercise 2
What is the total expected cost when the initial inventory is x = 0? What is the total
expected cost when the initial inventory x = 700? \


*\underline{Answer:}* \

Known that ordering cost is 1500 while c = 50, s = -15, sh = -100 with initial level x = 0,500,700,800.\
Therefore:\

\begin{eqnarray}
critical value &=& \frac{Cu}{Co+Cu} \nonumber\\
&=& \frac{50}{115}
\end{eqnarray}

with Theorem 2 and following uniform distribution therefore:\
\begin{eqnarray}
\frac{50}{115} &=& \frac{y^*-300}{900-300} \nonumber\\
y^* &=& 539.869 
\end{eqnarray}

Reading the problem we can calculate the total cost.\

\begin{eqnarray}
total cost = order cost + disruption cost + disposal cost
\end{eqnarray}

therefore when the initial inventory is x equal to 0 and the optimal value is 560.\
\begin{eqnarray}
\mathbb{E}[TC \textbar x = 0] &=& (1500 + 50\cdot X) + 100\cdot \mathbb{E}[(D-X)^+] + 15 \cdot \mathbb{E}[(X-D)^+]  \nonumber\\
&=& 26500 + 100 \cdot 96 + 15 \cdot 56 \nonumber\\
&=& 36940 
\end{eqnarray}

Then, the initial inventory is x equal to 700 and the optimal value is 560. But we know 700 > 560 so we should order 0\
\begin{eqnarray}
\mathbb{E}[TC \textbar x = 700] &=& (1500 + 50\cdot X) + 100\cdot \mathbb{E}[(D-X)^+] + 15 \cdot \mathbb{E}[(X-D)^+]  \nonumber\\
&=& 0 + 100 \cdot 600 + 0 \nonumber\\
&=& 60000
\end{eqnarray}

\newpage

## Exercise 3

Repeat the problem in page 16 when the demand is following:
$P(洧냥 = 300) = .2, P(洧냥 = 500) = .4, P(洧냥 = 700) = .3, P(洧냥 = 900) = .1$  fixed order cost became 1000.\

a. An optimal ordering.\

*\underline{Answer:}* \

\begin{eqnarray}
critical value &=& \frac{Cu}{Co+Cu} \nonumber\\
&=& \frac{50}{115}\nonumber\\
&=& 0.434
\end{eqnarray}

So based on the CDF given from the problem the optimal order is when D = 500 where the CDF is 0.6. Assuming the optimal order equal to demand then it will be equal to 500\

b. Expected for arbitrary initial inventory level.\

*\underline{Answer:}* \

\begin{eqnarray}
\mathbb{E}[TC \textbar order \: (500-x) \: if \: x=s] &=& \mathbb{E}[TC \textbar order \: 0 \: if \: x=s] \nonumber\\
1000 + 50 (500 - x) + 100 (500 - x) 15 (x - 500) &=& 0 + 100 \cdot 500 + 0 \nonumber\\
x &=& 137 
\end{eqnarray}

arbitary initial inventory level is 137.\

\newpage
c. Total expected cost when the initial inventory is x = 0 with the optimal ordering quantity 500 will be ordered .\

*\underline{Answer:}* \

\begin{eqnarray}
total cost = order cost + disruption cost + disposal cost
\end{eqnarray}

\begin{center}
\begin{tabular}{ c c c c}
 Demand & Total Cost in that Demand & Probability with Weight & Result \\
 D = 300 & $1000 + 50 \cdot 500 + 100\cdot 0 + 15 \cdot 200$ & $29000 \cdot 0.2$ & 5800\\ 
 D = 500 & $1000 + 50 \cdot 500 + 100\cdot 0 + 15 \cdot 0$ & $26000 \cdot 0.4$ & 10400\\ 
 D = 700 & $1000 + 50 \cdot 500 + 100\cdot 200 + 15 \cdot 0$ & $46000 \cdot 0.3$ & 13800\\
 D = 900 & $1000 + 50 \cdot 500 + 100\cdot 400 + 15 \cdot 0$ & $66000 \cdot 0.1$ & 6600  
\end{tabular}
\end{center}

therefore the expected cost is 36600.\

d. Total expected cost when the initial inventory is x = 700 with the optimal ordering quantity 500 will be ordered .\

*\underline{Answer:}* \

\begin{eqnarray}
total cost = order cost + disruption cost + disposal cost
\end{eqnarray}

\begin{center}
\begin{tabular}{ c c c c}
 Demand & Total Cost in that Demand & Probability with Weight & Result \\
 D = 300 & $1000 + 50 \cdot 500 + 100\cdot 0 + 15 \cdot 900$ & $39500 \cdot 0.2$ & 7900\\ 
 D = 500 & $1000 + 50 \cdot 500 + 100\cdot 0 + 15 \cdot 700$ & $36500 \cdot 0.4$ & 14600\\ 
 D = 700 & $1000 + 50 \cdot 500 + 100\cdot 200 + 15 \cdot 500$ & $33500 \cdot 0.3$ & 10050\\
 D = 900 & $1000 + 50 \cdot 500 + 100\cdot 400 + 15 \cdot 400$ & $32000 \cdot 0.1$ & 3200  
\end{tabular}
\end{center}

therefore the expected cost is 38950.\
 



\newpage
```{r, results='hide'}
"A7A8_Work"
```
